---
title: "Great Recession on micro and macro by Denny Li"
output: 
  flexdashboard::flex_dashboard:
      storyboard: true
      theme: yeti
  
---

```{r setup, include = FALSE, echo=FALSE}

if(!require(tidyverse)) install.packages("tidyverse")
library(tidyverse)

if(!require(DT)) install.packages("DT")
library(DT)

if(!require(flexdashboard)) install.packages("flexdashboard")
library(flexdashboard)

if(!require(lubridate)) install.packages("lubridate")
library(lubridate)

if(!require(scales)) install.packages("scales")
library(scales)

if(!require(maps)) install.packages("maps")
library(maps)

if(!require(CGPfunctions)) install.packages("CGPfunctions")
library(CGPfunctions)

knitr::opts_chunk$set(fig.width = 5, fig.asp = 1/3)


setwd("C:/Users/liden/Desktop/R final submission/R final submission")

load("Assignment.RData")

```


### 1. Median income {data-commentary-width=250}

```{r}

aChart <- ggplot(Income, aes(Year, `Median income`)) +
    geom_line(aes(color=State), size=1)+labs(title = "Median income between 2006 to 2010 by State")+geom_vline(aes(xintercept=as.numeric("2007")), lty=2)+
  geom_vline(aes(xintercept=as.numeric("2009")), lty=2)

  aChart    

plotly::ggplotly(aChart)

```
***
Median income slightly decreased from 2008 to 2009 in four states except Florida. Massachusetts had the highest median income and North Carolina had the lowest median income.  


### 2. Housing price by State
```{r}

housing_pricefivestates <- Housing %>% filter(RegionName %in% c("California", "Florida", "Massachusetts", "Michigan", "North Carolina"))


housing_pricefivestates$Date <- as.Date(housing_pricefivestates$Date, format="%m/%d/%Y")

housing_price <- housing_pricefivestates %>%
        group_by(RegionName, Year=year(Date)) %>%
  summarise(
       averageprice = mean(`Home Value Index (all homes)`))


housing_pricefilter <- housing_price %>% filter(Year > "2004" & Year < "2011")

 ggplot(housing_pricefilter, aes(x = Year, y = averageprice)) +
    geom_line(aes(color=RegionName), size=1)+labs(title = "Average home price by state")+geom_vline(aes(xintercept=as.numeric("2007")), lty=2)+geom_vline(aes(xintercept=as.numeric("2009")), lty=2)




 ggplot(housing_pricefilter, aes(x = Year, y = averageprice)) +
    geom_col(aes(fill=RegionName), position = "dodge")+labs(title = "Average home price by state")+geom_smooth()

 
 
states_map <- map_data("state")

state_map <- states_map %>% rename(RegionName=region)

housing_pricefilter <- housing_pricefilter %>% mutate(RegionName = fct_recode(RegionName, "california" = "California", "florida" = "Florida", "massachusetts" = "Massachusetts", "michigan" = "Michigan", "north carolina" = "North Carolina"))

housing_map <- inner_join(state_map, housing_pricefilter, by = "RegionName")

map_housing <- ggplot(housing_map, aes(x=long, y=lat, group=group, fill=averageprice))+
  geom_polygon(colour="black")+
  coord_map("polyconic")

map_housing

```

***
  The housing price in North Carolina remained relatively stable during the Great Recession (2007-2009) while other states saw significant decreases in home value. California had the highest average home price and Michigan had the lowest home price.  





### 3.Unemployment Rate
```{r}
Unemployment$DATE <- as.Date(Unemployment$DATE, format="%m/%d/%Y")

Unemployment_fivestates <- Unemployment %>%
        group_by(State, Year=year(DATE)) %>%
  summarise(avgUnemployRate = mean(`Unemployment Rate`))

ggplot(Unemployment_fivestates, aes(Year, avgUnemployRate)) +
    geom_area(aes(fill=State), size=2)+
    labs(title = "Unemployment Rate by state between 2006 to 2010")+geom_vline(aes(xintercept=as.numeric("2007")), lty=2)+
  geom_vline(aes(xintercept=as.numeric("2009")), lty=2)+
   theme(plot.title = element_text(size = 9))



ggplot(Unemployment_fivestates, aes(Year, avgUnemployRate)) +
    geom_line(aes(color = State), size=1)+
    labs(title = "Unemployment Rate by state between 2006 to 2010")+geom_vline(aes(xintercept=as.numeric("2007")), lty=2)+
  geom_vline(aes(xintercept=as.numeric("2009")), lty=2)+
   theme(plot.title = element_text(size = 9))


```



***
  The unemployment rate increased significantly across all five states during the Great Recession with four states (NC, MA, MI, CA) doubling and Florida tripling the unemployment rate. Michigan had the highest unemployment rate and Florida had the lowest unemployment rate. 




### 4. Crime Rate by State

```{r}
hChart <- ggplot(Crime_pivot, aes(Year, Rate)) +geom_area(aes(fill = State))+labs(title = "The Effect of Recession on Crime Rate")+
  facet_wrap(~`Crime types`, ncol = 2)

plotly::ggplotly(hChart)


```



  *** 
Surprisingly, the crime rate across all five states did not increase during the Great Recession. North Carolina had the highest burglary rate and Florida had the highest violent crime rate among five states. 
  
  
  

### 5. Depression thought 

```{r}

Depression <- Depression %>% 
  mutate(Age.Year = fct_recode(Age.Year, "18.2008-2009" = "X18....2008.2009.", 
"18.2014-2015" = "X18...2014.2015.", "12-17.2008-2009" = "X12.17..2008.2009.",
"12-17.2014-2015" = "X12.17..2014.2015.", "18-25.2008-2009" = "X18.25..2008.2009.", "18-25.2014-2015" = "X18.25..2014.2015.", 
"26.2008-2009" = "X26...2008.2009.", "26.2014-2015" = "X26...2014.2015."))

Depression <- Depression %>% rename(State = ï..State)  

Depression_fivestates <- Depression %>% filter(State %in% c("California", 
                  "Florida", "Massachusetts", "Michigan", "North Carolina"))



Depression_fivestates18 <- Depression_fivestates %>% 
  filter(Age.Year %in% c("18.2008-2009", "18.2014-2015"))


A <- newggslopegraph(Depression_fivestates18, 
                Age.Year, 
                Depression.Rate, 
                State, 
                Title = "Depression thought", 
                SubTitle = NULL, 
                Caption = NULL,
                LineThickness = 1.0,
                YTextSize = 3,
                LineColor = c(rep("gray",3), rep("gray",3), rep("gray",10)))

A


Depression_fivestates12_17 <- Depression_fivestates %>% 
  filter(Age.Year %in% c("12-17.2008-2009", "12-17.2014-2015"))


B <- newggslopegraph(Depression_fivestates12_17, 
                Age.Year, 
                Depression.Rate, 
                State, 
                Title = "Depression thought", 
                SubTitle = NULL, 
                Caption = NULL,
                LineThickness = 1.0,
                YTextSize = 3,
                )

B

Depression_fivestates18_25 <- Depression_fivestates %>% 
  filter(Age.Year %in% c("18-25.2008-2009", "18-25.2014-2015"))


C <- newggslopegraph(Depression_fivestates18_25, 
                Age.Year, 
                Depression.Rate, 
                State, 
                Title = "Depression thought", 
                SubTitle = NULL, 
                Caption = NULL,
                LineThickness = 1.0,
                YTextSize = 3,
                )

C
Depression_fivestates26 <- Depression_fivestates %>% 
  filter(Age.Year %in% c("26.2008-2009", "26.2014-2015"))

custom_colors <- tidyr::pivot_wider(Depression_fivestates26, 
                   id_cols = State, 
                   names_from = Age.Year, 
                   values_from = Depression.Rate) %>% 
  mutate(difference = 26.2014-2015 - 26.2008-2009) %>%
  mutate(trend = case_when(
    difference >= 0 ~ "green",
    difference < -1 ~ "red",
    TRUE ~ "gray"
    )
  ) %>%
  select(State, trend) %>%
  tibble::deframe()


D <- newggslopegraph(Depression_fivestates26, 
                Age.Year, 
                Depression.Rate, 
                State, 
                Title = "Depression thought", 
                SubTitle = NULL, 
                Caption = NULL,
                LineThickness = 1.0,
                YTextSize = 3,
                LineColor = c(rep("gray",3), rep("gray",3), rep("gray",10)))

D 





```

  
  ***
  The depression thought during the Great Recession (2008-2009) did not increase but decreasing among age groups 12-17 and 18-25 compared to non-recession period (2014-2015). While mixed results were found among 18+ and 26+ age groups with some states increasing and some states decreasing when two periods (2008-2009) and (2014-2015) were compared. 
  
  
  
### 6. Report card  

```{r}

Report_card1 <- Report_card %>% spread(Subject, Score)


Report_cardfilter <- Report_card1 %>% filter(Year <= "2011" & Year >= "2003")

states_map <- map_data("state")


states_map <- states_map %>% rename(State=region)

Report_cardfilter <- Report_cardfilter %>% mutate(State = fct_recode(State, "california" = "California", "florida" = "Florida", "massachusetts" = "Massachusetts", "michigan" = "Michigan", "north carolina" = "North Carolina"))

Report_card_map <- inner_join(states_map, Report_cardfilter, by="State")


Report_card_map <- Report_card_map %>% rename("Math Score"="Mathematics (scale range 0–500)")

Report_card_map <- Report_card_map %>% rename("Reading Score"="Reading (scale range 0–500)")




ggplot(Report_card_map, aes(State, `Math Score`)) +
    geom_col(aes(fill = as.factor(Year)), position = "dodge")+
    labs(title = "Math Score by state between 2003 to 2011")+
  theme(plot.title = element_text(size = 10),
        axis.text.x = element_text(size=7))+
  geom_hline(aes(yintercept=as.numeric("300")), lty=2)

ggplot(Report_card_map, aes(State, `Reading Score`)) +
    geom_col(aes(fill = as.factor(Year)), position = "dodge")+
    labs(title = "Reading Score by state between 2003 to 2011")+
  theme(plot.title = element_text(size = 10),
        axis.text.x = element_text(size=7))+
  geom_hline(aes(yintercept=as.numeric("275")), lty=2)



        

```


***
No significant changes in school performance, either Math or Reading, during the Great Recession. Massachusetts had the highest Math and Reading score.  

  
  
  